---
title: "Selecting Individuals and Loci"
output: 
  html_notebook:
    toc: true
---


## Intro/Overview

This notebook documents the strategy for 

1. selecting (typically at random) individuals to serve as Training Set individuals
to choose markers with large allele freq differences between farmed and wild salmon
1. then, once the training set is available, doing the actual locus selection in a way
that takes the most differentiated SNPs from each linkage block.
1. Finally creating data objects of the Test Set individuals with haplotypes
ready to simulate into offspring.

I had originally planned on doing this by using PLINK to quickly compute Fst's
for each locus, but I ultimately decided to keep it all in a tidy Hadley-esque
framework.  Accordingly, I will be selecting SNPs on the expected rate of correct
assignment of an allele to one group or the other---this seems to be a relevant measure
anyway, since what one is doing when inferring hybrids is assigning gene copies to one
group or the other.

```{r load-libs}
library(tidyverse)
library(stringr)

# and source the files that hold functions we will be using
source("R/assign_train_test.R")
source("R/compute_epp.R")
source("R/marker_selection.R")
```

## Training Set Selection

The AQU and WLN groups are the farmed fish in each data set.  So, we can get the data and put them
into different groups etc.  I am going to develop this with west and wrap it into a function later.
```{r get-data}
dat <- readRDS("intermediates/01/tidy-west.rds") %>%
  mutate(pop = str_sub(id, 1, 3),
         group = ifelse(pop == "AQU" | pop == "WLN", "farmed", "wild"))
```

This function assigns a fraction of fish to a training and a test group.
```r
`r paste(readLines("R/assign_train_test.R"), collapse = "\n")`
```


## Marker ranking and selection

### Computing the ranking criterion

We are going to rank markers on the basis of the expected posterior probability with which 
an allele would be assigned to the correct group.  Let $p_1$ and $p_2$ be the frequencies of
the $A$ allele at a SNP in two populations, 1 and 2.  The other allele, say $a$ has frequencies
of $1-p_1$ and $1-p_2$ in the populations.  If an allele is drawn and is type $A$, then 
the likelihood it came from population 1 is $p_1$ and from population 2 is $p_2$, and if we
had an equal prior on the origin of the gene copy the posterior that it came from population 1
would be $p_1/(p_1 + p_2)$.  So the expected posterior probability with which a gene copy from population 1
would be assigned to population 1 is simply:
\[
E_1 = \frac{p_1^2}{p_1 + p_2} + \frac{(1-p_1)^2}{2 - (p_1 + p_2)}
\]
and for a gene copy that was actually from population 2 we would have:
\[
E_2 = \frac{p_2^2}{p_1 + p_2} + \frac{(1-p_2)^2}{2 - (p_1 + p_2)}.
\]
So, let us propose as a statistic for ranking loci for their utility for NewHybrids analysis
what we call `epp`:
\[
\mathrm{epp} = \frac{E_1 + E_2}{2}
\]

To do this
we are going to need to 

1. compute the allele frequencies.  And we will add half of an allele to
each pile (essentially a unit information prior) when we do it.
1. compute the epp

Here is a function to do that:
```r
`r paste(readLines("R/compute_epp.R"), collapse = "\n")`
```

OK, that's that.   So, let's do it:
```{r do-epp}
set.seed(1234)  # for reproducibility during development
A <- assign_train_test(dat)
R <- compute_epp(A)
```

Something to note here is that I think the above calculations would also work for multiallelic
markers, like haplotypes. 

### Doing the marker selection

Once we have the ranks of the markers we need to do the selection of them.
In doing this we will want to be cognizant of physical linkage. I am not
going to screen things out on the basis of LD between loci.  My feeling is that
if there is LD within a single population between markers that are not very close 
physically (and hence will be filtered out anyway because of their physical proximity),
then those markers will likely be the result of recent strong selection in the aquaculture
strain (with very long blocks of LD).  In that case, the NewHybrids model is not quite correct, since it assumes
independence of parental population allele frequencies.  However, since we are simulating
things without any LD, I don't suspect the difference will be all that great.  The variance
in the posterior probabilities will be reduced a little because we are not simulating with 
LD, but I think that filtering on the basis of physical proximity is a better bet anyway.  

So, our method for marker selection will straightforward:

1. First, within each chromosome we rank all the variants by `epp` and then we include
the one with the highest epp in the chosen set. When a variant is placed in the chosen
set, we add its base-pair coordinates to a set $B$.
1. Then move on to the variant with the next highest `epp`.  If its base-pair coordinates
are greater than `min_dist` from any of the previously chosen variants in $B$, then it
will get chosen too, (and its coordindates included in $B$).

The above process continues until we have either selected (or not) all the loci on the
chromosome/linkage group.

Once all the linkage groups have been processed, we rank all the variants by their 
`epp` and select those that were on the "selected" pile from each linkage group.

The choice of `min_dist` in our case is determined by the sex-averaged map that we have.
After sex-averaging the recombination rates, our map changed in units of 0.05 cM, which 
corresponds to units of 50 Kb on the physical coordinates that we made.  This means
that any markers that are within 50 Kb of one another are essentially on the same
linkage group from the linkage map estimation project.  So, we will set min_dist to be 
60 Kb.  That will ensure that we don't take more than one marker from each group on the
sex-averaged linkage map.  

We will make a few functions for this here:
```r
`r paste(readLines("R/marker_selection.R"), collapse = "\n")`
```


And this is how we run it:
```{r run-m-select}
M <- marker_selection(R, min_dist = 6e04)
```

It would be interesting to look at those and see how many of them were on the original
linkage map, and how many were just "sprinkled in".  Here are the top 1000 loci
```{r show-M}
# get the sex-averaged map
sa <- readRDS("intermediates/01/sex_averaged_map.rds")

# then add a column "sprinkled"
M %>%
  mutate(sprinkled = !(variant %in% sa$SNP_ID))
```

## Save some files for the next segment

We are going to save the output of test_and_train and the marker selection:
```{r save-intermediates}
dir.create("intermediates/02", showWarnings = FALSE)
saveRDS(A, file = "intermediates/02/test_and_train.rds", compress = "xz")
saveRDS(M, file = "intermediates/02/marker_rankings.rds", compress = "xz")
```
